---
title: "Hitman-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hitman-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Hitman)
```

## Simulate data
To demonstrate the package on a dataset, I'll simulate log2 expression data with 100 genes and 6 samples in two groups with a continuous covariate and a binary outcome. The covariate and the first gene will be associated with group. This is similar to the example data from `limma` `contrasts.fit`.

We treat this data as being already processed. For many datasets, there are zeroes or missing values that might need to be imputed; samples need to be normalized to make them comparable and amenable to statistical tests; absent genes need to be removed; sample outliers need to be assessed to examine whether some experimental variables or batch effects need to be accounted for, or the samples need to be removed or down-weighted; and trends between a gene's mean expression and its variance should be accounted for, especially in RNA-seq data, with `limma`'s `voom` function.

```{r sim}
set.seed(42)
M <- matrix(rnorm(100*6, sd=0.3), nrow=100, ncol=6)
dimnames(M) <- list(paste0("gene", 1:nrow(M)), paste0("sample", 1:ncol(M)))
M[1,1:3] <- M[1,1:3] + 2

pheno <- data.frame(grp=rep(c("First3", "Last3"), each=3), yy=rnorm(ncol(M)), stringsAsFactors = FALSE)
pheno$yy[1:3] <- pheno$yy[1:3]-2
```

## Hitman
We  can test what genes mediate the effect of group on the continuous phenotype. This is called high-throughput mediation analysis (`hitman`). We set the exposure (`E`) to group, the outcome (`Y`) to `pheno.v`, and test each gene in `M` as a mediator. For `hitman` we transform the exposure to numeric, and set the names of the `phenoype` to those of `M`.  
```{r}
exposure <- as.numeric(pheno$grp == "Last3")
res.hit <- hitman(E = exposure, M = M, Y = setNames(pheno$yy, nm=colnames(M)))
```

`hitman` returns a table with a combined ("EMY") p-value and FDR for each gene. This combined statistic is based on the association of E to the gene in `M`, and the conditional association of the gene in `M` to `Y` given `E`.
```{r, echo=FALSE}
knitr::kable(signif(res.hit[1:10,], 3))
```
